---
title: "Klassification"
author: "Anders Thor Kellmer"
date: "2023-06-12"
output: html_document
---

## The iris  data
```{r, warning=FALSE, message=FALSE}
library(MASS)
library(caret)
set.seed(123)
```

The iris data set is a classical data set first studied by R. A. Fisher in 1936. It contains measurements of sepal and petal length and width for three different species of the iris plant.

Load the data set. Rename the species to 1, 2 and 3 for simplicity.

```{r load data}
data(iris)
str(iris)
X = iris[, 1:2]
y = rep(0, nrow(iris)) 
y[iris[, 5] == "setosa"] = 1
y[iris[, 5] == "versicolor"] = 2
y[iris[, 5] == "virginica"] = 3
niris = data.frame(Species = y,
                   SLength = X[, 1],
                   SWidth = X[, 2])
```

Randomly split the data set.

```{r split data}
dt = sort(sample(nrow(niris), nrow(niris)*0.80))
train = niris[dt, ]
test = niris[-dt, ]
```

Fit the LDA and QDA models on the training set.

```{r fit}
lda.fit = lda(Species ~ ., niris)
qda.fit = qda(Species ~ ., niris);
lda.fit
qda.fit
```

Predict on the test set using the fitted models.

```{r predict}
new.pred.lda = predict(lda.fit, test)
new.pred.qda = predict(qda.fit, test)
```

Calculate confusion matrix and missclassification error.

```{r missclassification}

confusionMatrix(as.factor(test[[1]]), as.factor(new.pred.lda$class))$table
confusionMatrix(as.factor(test[[1]]), as.factor(new.pred.qda$class))$table

mean(ifelse(test[, 1] == new.pred.lda$class, 0, 1))
mean(ifelse(test[, 1] == new.pred.qda$class, 0, 1))
```

Generate a grid on a square that contains all observations.

```{r grid}
xmin = min(niris$SLength) - 1
xmax = max(niris$SLength) + 1
ymin = min(niris$SWidth) - 1
ymax = max(niris$SWidth) + 1

n_points = 150     
x_grid = seq(xmin,xmax,length.out = n_points) 
y_grid = seq(ymin,ymax,length.out = n_points) 
```

Plot the predicted classes along with the grid using LDA.
```{r lda plot}
par(mar = c(0.1, 0.1, 2, 0.1))
plot(
  X,
  pch = y - 1,
  col = y + 1,
  main = "LDA",
  yaxt = 'n',
  xaxt = 'n'
)

for (j in 1:n_points) {
  for (k in 1:n_points) {
    new = data.frame(SLength = x_grid[j], SWidth = y_grid[k])
    new.pred.lda = predict(lda.fit, new)
    curr_obs = as.numeric(new.pred.lda$class)
    points(x_grid[j],
           y_grid[k],
           col = curr_obs + 1,
           pch = 3,
           cex = 0.3)
  }
}
# Red = Setosa, Green = Versicolor, Blue = Virginica
```

Plot the predicted classes along with the grid using QDA, where red is Setosa, green is Versicolor and blue is Virginica. 

```{r qda plot}
par(mar = c(0.1, 0.1, 2, 0.1))
plot(
  X,
  pch = y - 1,
  col = y + 1,
  main = "QDA",
  yaxt = 'n',
  xaxt = 'n'
)

for (j in 1:n_points) {
  for (k in 1:n_points) {
    new = data.frame(SLength = x_grid[j], SWidth = y_grid[k])
    new.pred.qda = predict(qda.fit, new)
    curr_obs = as.numeric(new.pred.qda$class)
    points(x_grid[j],
           y_grid[k],
           col = curr_obs + 1,
           pch = 3,
           cex = 0.3)
  }
}
# Red = Setosa, Green = Versicolor, Blue = Virginica
```

