---
title: "5. Support Vector Machines"
author: "Anders Thor Kellmer"
date: "2023-06-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

Load dependencies.

```{r dependencies, message=FALSE}
library("e1071")
library(GGally)
library(ggplot2)
library(caret)
```

Exploratory analysis.

```{r data import, message=FALSE}
data("iris")
head(iris, n = 5)
ggpairs(iris, ggplot2::aes(colour = Species, alpha = 0.4))
```

### Linear SVM

Fitting linear SVM on iris dataset.

```{r model}
svm_model <- svm(
  Species ~ ., 
  data = iris,
  kernel = "linear"
)
```

Plotting model in two dimensions using **Petal.Width** and **Petal.Length**.

```{r}
plot(
  svm_model,
  data = iris,
  Petal.Width ~ Petal.Length,
  slice = list(Sepal.Width = 3, Sepal.Length = 4)
)
```

Calculate confusion matrix and missclassification rate.

```{r}
pred = predict(svm_model, iris)
confusionMatrix(as.factor(iris[[5]]), as.factor(pred))$tabl
mean(ifelse(iris[[5]] == pred, 0, 1))
```

### Radial SVM

Fitting radial SVM on iris dataset.

```{r}
svm_model = svm(
  Species ~ ., 
  data=iris,
  kernel = "radial"
)
```

Plotting model in two dimensions using **Petal.Width** and **Petal.Length**.

```{r}
plot(
  svm_model, 
  data = iris,
  Petal.Width ~ Petal.Length,
  slice = list(Sepal.Width = 3, Sepal.Length = 4) 
)
```

Calculate confusion matrix and missclassification rate.

```{r}
pred = predict(svm_model, iris)
confusionMatrix(as.factor(iris[[5]]), as.factor(pred))$tabl
mean(ifelse(iris[[5]] == pred, 0, 1))
```

### Tuning SVM
```{r}
tmodel = tune(
  svm,
  Species~., 
  data = iris,
  ranges = list(epsilon = seq(0,1,0.1), cost = 2^(2:7))
)
best_model <- tmodel$best.model
best_model
```

Plot of the best model.

```{r}
plot(
  best_model, 
  data = iris,
  Petal.Width ~ Petal.Length,
  slice = list(Sepal.Width = 3, Sepal.Length = 4)  
)
```

Confusion matrix and missclassification rate.

```{r}
pred_best_model = predict(best_model,iris)
confusionMatrix(as.factor(iris[[5]]), as.factor(pred_best_model))$tabl
mean(ifelse(iris[[5]] == pred_best_model, 0, 1))
```

